Joint Power Allocation and Network Slicing in an End-to-End ORAN System \vspace{-.1cm}

### 2019-10-29 Latex/tex/pr.tex
last version -revised

\begin{abstract}
Many major telecommunication companies confirmed the unification of the xRAN Group with the C-RAN Alliance to establish a more flexible and openness radio access network which is the Open-RAN
(ORAN) for the fifth generation of wireless technology. To increase energy efficiency and optimize the allocation of resources, Network Slicing (NS) is considered as the best method for the fifth generation (5G) in order to virtualize the common physical network into several logical end-to-end networks. Every slice consists of a part of core network resources, network functions, and radio access network resources as a functional end-to-end network. In this paper, we elaborate joint NS in RAN and Core of ORAN system to investigate the power
of each User Equipment (UE), map slices to services and also map physical Data Centers (DC) to slices to jointly maximize energy efficiency and minimize consumption power of RUs and the cost of  physical resources in a downlink channel. The problem is formulated as a mixed-integer optimization problem that can be decomposed into two independent sub-problems due to the fact that sub-problems are independent.
Heuristic algorithms are proposed to each of sub-problems in order to map slices to services, optimize power consumption and map slices to physical resources to minimize the cost of total DCs simultaneously.
\end{abstract}

@@

ORAN virtualizes the elements of the radio access networks, separate them and define appropriate open interfaces for connection of these elements. Moreover, ORAN uses machine learning techniques to develop smarter RAN layers in its architecture. In an innovative ORAN system, the programmable RAN software is decoupled from hardware \cite{oran1}. Open interface is one of the most crucial properties for ORAN to enable mobile network operators (MNOs) to define their own services.
The concept of software defined network (SDN), which is the separation of control plane from user plane, is deployed in an intelligent ORAN architecture. Moreover, this separation promotes RRM to use non-realtime (RT) and near-realtime (NRT) RAN Intelligent Controller (RIC).
In the ORAN architecture, distributed unit (DU) is a logical node hosting RLC/MAC/High-PHY layer.
Moreover, central unit (CU) is a logical node hosting RRC, SDAP and PDCP. Also, radio unit (RU)  is a logical node hosting LOW-PHY layer and
RF processing \cite{oranpaper}.
ORAN introduces interfaces such as open fronthaul interface that connects DU and RU(i.e., E2 interface), and an A1 interface between orchestration/NMS layer containing the non-real-time RIC (RIC non-RT) function and the eNB/gNB containing the near-real-time RIC (RIC near-RT) function.

To evolve servicing in 5G networks, separation of elements of software and hardware of network is employed and is referred to as network functions virtualization (NFV). Virtual network function (VNF) are functional blocks of the system in this case. 5G networks are expected to host several services with different requirements simultaneously. Network slicing is considered as a solution for such demand. A network slice is a logical end-to-end network which provide service with specific requirements. Multiple network slices run on the same physical network infrastructure which are managed and operate independently \cite{lee2018dynamic}. Slicing can be defined for the RAN, for the core network or both of these networks \cite{ns1}.

@@

In this paper, as  depicted in Figure \ref{fig:c11}, the downlink of the ORAN system is studied. The RAN is divided to three layers including RU, DU, and CU with open interfaces. DU and CU are running on general purpose data centers. UEs are divided to different services according to their service requirements. RAN resources are decoupled to slices to provide requirements of services. Optimal power allocation and joint mapping of slices to services are applied. In addition, mapping slices to physical resources is taken to account. The contribution of the paper is expressed as follow:
\begin{itemize}
\item In this paper, joint network slicing and resource allocation is considered in an end to end ORAN system.
\item UEs are categorized into different class of services based on their service requirements. RAN which contains RUs, PRBs, and VNFs ,in the two layer of processing, are classified into different group of slices.
\item The problem is decomposed into two independent sub-problems.
\item Heuristic algorithms are applied for the sub-problems to efficiently obtain the solution.
\end{itemize}

@@

single-antenna user equipments (UEs) that require certain service. Each slice $s \in \{1,2,...,S \}$ consists of $R_s$ RUs and $K_s$ physical resource blocks (PRBs), one DU and one CU that contains VNFs. All RUs in a slice, that is mapped to a service, transmit signals cooperatively to all the UEs in a specific service \cite{motalleb2017optimal,mimoCran}. Each RU $r \in \{1,2,...,R \}$ is mapped to a DU via an optical fiber link with limited fronthaul capacity.
There are two processing layers one in the DU and one on the CU of ORAN system, each represented with a VNF. The lower layer (i.e., DU) consists of high-PHY, MAC, and RLC, and the upper layer (i.e., CU) consists of RRC, PDCP and SDAP. Assume we have $M_1$ VNFs in the DU layer and $M_2$ VNFs in the CU layer for processing data.

@@

where $1/\mu_1$ and $1/\mu_2$ are the mean service time of the first and the second layers respectively.%, due to the fact that the service time of each data packet has an exponential distribution for $\mu_i>\alpha_{s}/{M_{s,i}}$ $ i \in \{1,2\}$. 

@@

 is equal to the arrival data rate of dispatchers (used as a traffic divider) for each slice \cite{frdl}.

@@

the physical DC is connected to a VNFs of a slice or not.

@@

ensures that each service is connected to at least one slice.

@@

The optimization problem in \eqref{constraints} can be decomposed into two independent optimization problems A and B since the variables can be obtained independently and respectively. In problem A,
the variables $\boldsymbol{P}$ and $ \boldsymbol{A}$ is obtained. The problem A is as follow

@@

and $\eta$, the two parts of the problems are solved. We repeat this procedure until the algorithm is converged.

@@

\State Sort slices according to the weighted linear combination of number of PRBs, RUs and VNFs in two layers and the Capacity of their resources in the descending order.

@@

\begin{equation}\label{lagrang}

@@

&+ \sum\limits_{v=1}^{V} \sum\limits_{k=1}^{U_v} \kappa_{u_{(v,k)}} \sum\limits_{s=1}^{S}(R_{u_{(v,k)}} -\mathfrak{D_s})a_{v,s}.\\

@@

Firstly, we obtain $\boldsymbol{A}$ by fixing $\boldsymbol{P} = P_{max}$ in the problem \eqref{constraints1} and using algorithm \eqref{alg}. Also we fixed $\eta = 0$. Afterward, by achieving $\boldsymbol{A}$, in the second part, we acquire $\boldsymbol{P}$ using the sub-gradient method. After solving $\boldsymbol{P}$, $\eta$ is updated. Then in the next iteration, with new $\boldsymbol{P}$
and $\eta$ the two parts of the problems is solved until the algorithm converged.

@@

Secondly, we start mapping from the most needed slices to the DC with the most physical resources (from line \ref{33} to line \ref{34} of algorithm \ref{alg3}). After mapping DCs to slices, if some slices are not admitted, we start admitting remaining slices to more than one DCs (from line \ref{35} to line \ref{36} of algorithm \ref{alg3}). At the end, if DC with the lowest physical resources is free and can be served instead of DC with the highest physical resource, the slices remapped to new DC with the lowest physical resource since it has the lowest power consumption (line \ref{37} of algorithm \ref{alg3}).
If a slice does not admit to a specific DC, it remains for the next placement. In next placement the remaining slices, map to more than one DC according to their requirements.

@@

In Fig. \ref{fig:f1}, the ratio of admitted slices is demonstrated for two different numbers of DCs with the different number of slices (the parameters for simulation listed in table \ref{table:1} and also we set $w_C = 320$, $w_S = 100$, $w_M =1$). In this simulation, it is assumed that  just one DC can serve each slice, and it is not admitted by more than one DC. The proposed method is based on Algorithm \ref{alg3}, and the optimal method is done by the MOSEK toolbox.

@@

In Fig. \ref{fig:f2}, the normalized resource consumption is depicted due to the number of slices (the parameters for simulation listed in table \ref{table:1} and also we set $w_C = 320$, $w_S = 100$, $w_M =1$). In this simulation, it is assumed that the number of DCs is entirely enough to cover all slices. The optimality of the placement of DCs to slices is measured. It is shown that how much resources of active DCs are not used. For ten slices, the difference between the optimal solution and the proposed solution is about $15$ percent.

>>>>

\begin{abstract}
Many major telecommunication companies confirmed the unification of the xRAN Group with the C-RAN Alliance to establish a more flexible and openness radio access network which is the Open-RAN
(ORAN) for the fifth generation of wireless technology. To increase energy efficiency and optimize the allocation of resources, Network Slicing (NS) is considered as the best method for the fifth generation (5G) in order to virtualize the common physical network into several logical end-to-end networks. Every slice consists of a part of core network resources, network functions, and radio access network resources as a functional end-to-end network. In this paper, we elaborate joint NS in RAN and Core of ORAN system to investigate the power
of each User Equipment (UE), map slices to services and also map physical Data Centers (DC) to slices to jointly maximize energy efficiency and minimize consumption power of RUs and the cost of  physical resources in a downlink channel. The problem is formulated as a mixed-integer optimization problem that can be decomposed into two independent sub-problems.
Heuristic algorithms are proposed to each of sub-problems in order to map slices to services, optimize power consumption and map slices to physical resources to minimize the cost of total DCs simultaneously.
\end{abstract}

@@

ORAN virtualizes the elements of the radio access networks, separate them and define appropriate open interfaces for connection of these elements. Moreover, ORAN uses machine learning techniques to develop smarter RAN layers in its architecture. In an innovative ORAN system, the programmable RAN software is decoupled from hardware \cite{oran1}. Open interface is one of the most crucial properties for ORAN to enable mobile network operators (MNOs) to define their own services.
The concept of software defined network (SDN), which is the separation of control plane from user plane, is deployed in an intelligent ORAN architecture. Moreover, this separation promotes RRM to use non-realtime (RT) and near-realtime (NRT) RAN Intelligent Controller (RIC).
In the ORAN architecture, distributed unit (DU) is a logical node having RLC/MAC/High-PHY layer.
Moreover, central unit (CU) is a logical node having RRC, SDAP and PDCP. Also, radio unit (RU) is a logical node having LOW-PHY layer and
RF processing \cite{oranpaper}.
ORAN introduces interfaces such as open fronthaul interface that connects DU and RU (i.e., E2 interface), and an A1 interface between orchestration/NMS layer containing the non-real-time RIC (RIC non-RT) function and the eNB/gNB containing the near-real-time RIC (RIC near-RT) function.

To evolve servicing in 5G networks, separation of elements of software and hardware of network is employed and referred as network functions virtualization (NFV). Virtual network function (VNF) are functional blocks of the system in this case. 5G networks are expected to host several services with different requirements simultaneously. Network slicing is considered as a solution for such demand. A network slice is a logical end-to-end network which provide service with specific requirements. Multiple network slices run on the same physical network infrastructure which are managed and operate independently \cite{lee2018dynamic}. Slicing can be defined for the RAN, for the core network or both of these networks \cite{ns1}.

@@

In this paper, as  depicted in Figure \ref{fig:c11}, the downlink of the ORAN system is studied. The RAN is divided into three layers including RU, DU, and CU with open interfaces. DU and CU are running on general purpose data centers. UEs are divided to different services according to their service requirements. RAN resources are decoupled to slices to provide requirements of services. Optimal power allocation and joint mapping of slices to services are applied. In addition, mapping slices to physical resources is taken to account. The contribution of the paper is expressed as follow:
\begin{itemize}
\item In this paper, joint network slicing and resource allocation is considered in an end to end ORAN system.
\item UEs are categorized into different class of services based on their service requirements. RAN which contains RUs, PRBs, and VNFs ,in the two layer of processing, are classified into different group of slices.
\item The problem is decomposed into two independent sub-problems.
\item Heuristic algorithms are applied for the sub-problems to efficiently obtain the solution.
\end{itemize}

@@

single-antenna user equipments (UEs) that require certain service. Each slice $s \in \{1,2,...,S \}$ consists of $R_s$ RUs and $K_s$ physical resource blocks (PRBs), one DU and one CU that contains VNFs.
Slices can have shared resources. All RUs in a slice, that is mapped to a service, transmit signals cooperatively to all the UEs in a specific service \cite{motalleb2017optimal,mimoCran}. Each RU $r \in \{1,2,...,R \}$ is mapped to a DU via an optical fiber link with limited fronthaul capacity.
There are two processing layers one in the DU and one in the CU of ORAN system, each represented with a VNF. The lower layer (i.e., DU) consists of high-PHY, MAC, and RLC, and the upper layer (i.e., CU) consists of RRC, PDCP and SDAP. Assume we have $M_1$ VNFs in the DU layer and $M_2$ VNFs in the CU layer for processing data.

@@

where $1/\mu_1$ and $1/\mu_2$ are the mean service time of the first and the second layers respectively.

@@

 is equal to the arrival data rate of load balancers for each slice \cite{frdl}.

@@

the physical DC is mapped to a VNFs of a slice or not.

@@

ensures that each service is mapped to at least one slice.

@@

The optimization problem in \eqref{constraints} can be decomposed into two independent optimization problems A and B since the variables can be obtained independently and respectively. Firstly we need to solve problem A. After obtaining $\boldsymbol{P}$ and $ \boldsymbol{A}$, problem B can be solved by having the value of $ \boldsymbol{A}$. 
The problem A is as follow

@@

and $\eta$, the two parts of the problems are solved. We repeat this procedure until the algorithm  converges.

@@

\State Sort slices according to the weighted linear combination of number of PRBs, RUs and VNFs in two layers and the capacity of their resources in the descending order.

@@

\begin{subequations}\label{lagrang}

@@

&+ \sum\limits_{v=1}^{V} \sum\limits_{k=1}^{U_v} \kappa_{u_{(v,k)}} \sum\limits_{s=1}^{S}(R_{u_{(v,k)}} -\mathfrak{D_s})a_{v,s}.

@@

Firstly, we obtain $\boldsymbol{A}$ by fixing $\boldsymbol{P} = P_{max}$ in the problem \eqref{constraints1} and using algorithm \eqref{alg}. Also we fixed $\eta = 0$. Afterward, $\boldsymbol{A}$ is achieved from algorithm \ref{alg}. Then, in the second part, we acquire $\boldsymbol{P}$ using the sub-gradient method. After solving $\boldsymbol{P}$, $\eta$ is updated. Then in the next iteration, with new $\boldsymbol{P}$
and $\eta$ the two parts of the problems is solved until the algorithm converges.

@@

Secondly, we start mapping from the most needed slices to the DC with the most physical resources (from line \ref{33} to line \ref{34} of algorithm \ref{alg3}). After mapping DCs to slices, If a slice does not admit to a specific DC, it remains for the next placement. In the next placement the remaining slices, map to more than one DC according to their requirements. (from line \ref{35} to line \ref{36} of algorithm \ref{alg3}). At the end, if DC with the lowest physical resources is free and can be served the slices of a DC with the highest physical resource, the slices remapped to new DC with the lowest physical resource since it has the lowest power consumption (line \ref{37} of algorithm \ref{alg3}).

@@

In Fig. \ref{fig:f1}, the ratio of admitted slices is demonstrated for two different numbers of DCs with the different number of slices (the parameters for simulation listed in table \ref{table:1} and also we set $w_C = 320$, $w_S = 100$, $w_M =1$). In this simulation, the second term of \eqref{eqpsi}, is important and the designed parameter $\nu$ is supposed to be high.
Also, it is assumed that just one DC can serve each slice, and each slice is not admitted by more than one DC. 
The proposed method is based on Algorithm \ref{alg3}, and the optimal method is done by the MOSEK toolbox.

@@

In Fig. \ref{fig:f2}, the normalized resource consumption is depicted due to the number of slices (the parameters for simulation listed in table \ref{table:1} and also we set $w_C = 320$, $w_S = 100$, $w_M =1$). In this simulation, suppose the number of DCs is entirely enough to cover all slices and also, the designed parameter $\nu$ is assumed to be low, so we focused on the first term of \eqref{eqpsi}. However, if any slice remains, it can be served by more than one DCs. The optimality of the placement of DCs to slices is measured based on the power consumption of DCs. It is shown that how much resources of active DCs are not used. For ten slices, the difference between the optimal solution and the proposed solution is about $15$ percent.
